On the entropy of isothermal dusty gas: application to protoplanetary disks

### 2016-08-01 appendix.tex
general edits
regular update

\section{Linearized dust diffusion}\label{lin_dust}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop  =  \frac{K}{\rho c_s}
\end{align}
\citep{price15}, where $K$ is a constant. Then the dust diffusion
function becomes
\begin{align}
  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\epsilon\tstop\nabla
  P\right) = - K c_s^2 \nabla\cdot
  \left(\frac{\epsilon}{c_s}\bm{F}\right) =
  -Kc_s \left(\bm{F}\cdot\nabla\epsilon + \epsilon \nabla\cdot\bm{F}
  - \frac{1}{2}\epsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
\end{align}
Linearizing,
\begin{align}
  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\epsilon +
  \dd\bm{F}\cdot\nabla\epsilon + \dd\epsilon\nabla\cdot\bm{F} +
  \epsilon\nabla\cdot\dd\bm{F} -
  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\epsilon + \epsilon
  \dd\bm{F}\right). 
\end{align}
The linearized dust-fraction $\dd\epsilon$ and its derivatives in the
radially-local approximation are given by
\begin{align}
  \dd\epsilon      &= (1-\epsilon)W - \frac{Q}{c_s^2}, \\
  \p_r\dd\epsilon &= \left[(1-\epsilon)\left(\ii k -
    \p_r\ln{\rho}\right) - \p_r\epsilon\right]W + \left(\p_r\ln{c_s^2}
  + \p_r\ln{\rho} - \ii k \right)\frac{Q}{c_s^2},\\
  \dd\epsilon^\prime &= (1-\epsilon)W^\prime - \epsilon^\prime W -
  \left(\frac{Q}{c_s^2}\right)^\prime. 
\end{align}

\section{Conversion formulae}
The gas density $\rhog$ and dust-to-gas ratio $\tepsilon$ is related
to the total density $\rho$ and dust-fraction $\epsilon$ by
\begin{align}
  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \tepsilon\right)},\\
  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\tepsilon}{1+
    \tepsilon},\\
  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
  \frac{\nabla^2\tepsilon}{1+\tepsilon} -
  \frac{\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^2}, 
\end{align}
and
\begin{align}
  \epsilon &= \frac{\tepsilon}{1+ \tepsilon},\\
 \nabla\epsilon & =
 \frac{\nabla\tepsilon}{\left(1+\tepsilon\right)^2},\\
 \nabla^2\epsilon &=
 \frac{\nabla^2\tepsilon}{\left(1+\tepsilon\right)^2} -
 \frac{2\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^3}. 
\end{align}

>>>>

\section{Linearized dust diffusion}\label{lin_dust}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop  =  \frac{K}{\rho c_s}
\end{align}
\citep{price15}, where $K$ is a constant. Then the dust diffusion
function becomes
\begin{align}
  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right) = - K c_s^2 \nabla\cdot
  \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
  -Kc_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
  - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
\end{align}
Linearizing,
\begin{align}
  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\tepsilon +
  \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
  \tepsilon\nabla\cdot\dd\bm{F} -
  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\tepsilon + \tepsilon
  \dd\bm{F}\right). 
\end{align}
The linearized dust-fraction $\dd\tepsilon$ and its derivatives in the
radially-local approximation are given by
\begin{align}
  \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k -
    \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
  + \p_r\ln{\rho} - \ii k \right)\frac{Q}{c_s^2},\\
  \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
  \left(\frac{Q}{c_s^2}\right)^\prime. 
\end{align}

\section{Conversion formulae}
The gas density $\rhog$ and dust-to-gas ratio $\epsilon$ is related
to the total density $\rho$ and dust-fraction $\tepsilon$ by
\begin{align}
  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \epsilon\right)},\\
  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\epsilon}{1+
    \epsilon},\\
  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
  \frac{\nabla^2\epsilon}{1+\epsilon} -
  \frac{\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^2}, 
\end{align}
and
\begin{align}
  \tepsilon &= \frac{\epsilon}{1+ \epsilon},\\
 \nabla\tepsilon & =
 \frac{\nabla\epsilon}{\left(1+\epsilon\right)^2},\\
 \nabla^2\tepsilon &=
 \frac{\nabla^2\epsilon}{\left(1+\epsilon\right)^2} -
 \frac{2\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^3}. 
\end{align}

### 2016-08-01 linear.tex
regular update

\section{Linear problem for perfectly coupled dust}\label{linear_problem}
In this paper we calculate explicit solutions to the linearized
equations in the limit of perfectly coupled dust ($\tstop=0$), so that
the equilibria defined in \S\ref{eqm} are exact steady states; and we
consider radially-localized axisymmetric disturbances of the form  
\begin{align}
  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k r)},
\end{align} 
%and similarly for $\dd P$ and $\dd\bm{v}$. 
where $k$ is a real wavenumber such that $|kr|\gg 1$, and the
amplitude $\dd f_1(r,z)$ is 
a slowly-varying function of $r$. Then 
$\p_r\to i k$ when acting on the above primitive perturbations, and we may
neglect curvature terms. We take  
$k>0$ without loss of generality. Hereafter, we drop the subscript 1
on the amplitudes. 
%The frequency $\sigma = \omega +
%\ii s$ is generally complex, with $\omega$ being the real frequency
%and $s$ is the real growth rate.

@@

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All coefficients and amplitudes are
evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
is retained.  Two boundary conditions are needed for the $\tstop=0$
problem. This is expected since this problem is exactly equivalent to
that in adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$ and
$\zmax=5\Hgas$.

@@

%Now, for finite dust-gas coupling, $\tstop\neq0$, inspection of $\dd C$
%(Appendix \ref{lin_dust}) shows that it involves $W^\prime$ (assuming
%$\epsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
%differential equation for $W$. However, for perfectly coupled dust,
%$\tstop =\dd\mathcal{C}= 0$. In that case Eq. \ref{lin_energy} gives an algebraic
%relation between $W$, $Q$ and $\dd v_z$. That is,
%\begin{align*}
%\text{Eq. \ref{lin_energy}}\Rightarrow \begin{cases}
%  W^\prime(z) = W^\prime(W, Q, \dd v_z) & \text{if } \tstop \neq 0, \\
%  W           (z) = W(Q, \dd v_z) & \text{if }\tstop = 0.
%\end{cases}
%\end{align*}

>>>>

\section{Linear problem for perfectly coupled dust}\label{linear_problem}
We calculate explicit solutions to the linearized
equations in the limit of perfectly coupled dust ($\tstop=0$), so that
the equilibria defined in \S\ref{eqm} are exact steady states. We 
consider radially-localized axisymmetric disturbances of the form  
\begin{align}
  \delta f (r, z) = \delta f_1(r,z)\exp{(\ii k r)},
\end{align} 
%and similarly for $\dd P$ and $\dd\bm{v}$. 
where $k$ is a real wavenumber such that $|kr|\gg 1$, and the
amplitude $\dd f_1(r,z)$ is 
a slowly-varying function of $r$. Then 
$\p_r\to i k$ when acting on the above primitive perturbations, and we may
neglect curvature terms. We take  
$k>0$ without loss of generality. Hereafter, we drop the subscript 1
on the amplitudes. 
%The frequency $\sigma = \omega +
%\ii s$ is generally complex, with $\omega$ being the real frequency
%and $s$ is the real growth rate.

@@

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All coefficients and amplitudes are
evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
is retained.  Two boundary conditions are needed for the $\tstop=0$
problem. This is expected since this problem is exactly equivalent to
 adiabatic hydrodynamics \citep[e.g.][]{lubow93}. For simplicity
we impose solid boundaries so that $\delta v_z(\pm\zmax)=0$ and
$\zmax=5\Hgas$.

@@

%Now, for finite dust-gas coupling, $\tstop\neq0$, inspection of $\dd C$
%(Appendix \ref{lin_dust}) shows that it involves $W^\prime$ (assuming
%$\tepsilon < 1$). Then Eq. \ref{lin_energy} may be taken as a
%differential equation for $W$. However, for perfectly coupled dust,
%$\tstop =\dd\mathcal{C}= 0$. In that case Eq. \ref{lin_energy} gives an algebraic
%relation between $W$, $Q$ and $\dd v_z$. That is,
%\begin{align*}
%\text{Eq. \ref{lin_energy}}\Rightarrow \begin{cases}
%  W^\prime(z) = W^\prime(W, Q, \dd v_z) & \text{if } \tstop \neq 0, \\
%  W           (z) = W(Q, \dd v_z) & \text{if }\tstop = 0.
%\end{cases}
%\end{align*}
